\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-28 Draft_Form/tex/Materials_and_Methods(2).tex
update 28_02_2018
update 28_02_2018

The German NFI is a periodic inventory that is carried out every 10 years over the entire forest area of Germany. The most recent inventory (BWI3) was conducted in 2011 and 2012. While information was originally gathered on a systematic 4x4 km grid, some federal states such as Rhineland-Palatinate have switched to a densified 2x2 km grid. The German NFI uses a cluster sampling design, which means that a sample unit consists of at most four sample locations (also referred to as \textit{sample plots}) that are arranged in a square, called \textit{cluster}), with a side length of 150 metres. The number of plots per cluster can vary between 1 and 4 depending on forest/non-forest decisions by the field crews on the individual plot level \citep{bwi3_aufn}. In the field survey of the BWI3, sample trees for timber volume estimation are selected according to the angle count sampling technique \citep{bitterlich1984}, using a basal area factor ($BAF$) of 4 that is respectively adjusted for sample trees at the forest boundary by a geometric intersection of the boundary transect with the tree-individual inclusion circle \citep{bwi3_aufn}. A further inventory threshold for a tree to be recorded is a diameter at breast height ($DBH$) of at least 7 cm. For each sample tree that is selected by this procedure, the DBH, the absolute tree height, the tree diameter at 7 m ($D7$) and the tree species is measured and used to estimate the volume at the tree level. These volume estimates are based on the application of tree species specific taper curves that are adjusted to the set of diameters and corresponding height measurements taken from the respective sample tree \citep{kublin2013}.

@@

The estimators used in this study have been proposed by \citep{mandallaz2013a,mandallaz2013b} and derive their mathematical properties under the so-called infinite population approach (IPA). Therefore, we shall first provide a short introduction into this general estimation framework. We start by assuming that the population $P$ of trees $i \in 1,2, ..., N$ within a forest of interest $F$ is exactly defined, and each tree $i$ has a response variable $Y_i$ (e.g. its timber volume) that can be used to define the population mean $Y$ (e.g., the average timber volume per unit area) over $F$. Since a full census of all tree population individuals is almost never feasible, $Y$ has to be estimated based on a sample. In the infinite population approach this sample is a set of points or locations $x$ distributed independently and uniformly over the set of all possible points in $F$. Each point $x$ has an associated local density $Y(x)$ (e.g., the timber volume per unit area) whose spatial distribution is given by a fixed (i.e. non stochastic) piecewise constant function. The population mean $Y$ is mathematically equivalent to the integral of the local density function surface divided by the surface area of $F$, $\lambda(F)$, i.e. $Y=\frac{1}{N} \sum_{i=1}^{N}Y_i=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$, and thus the population mean $Y$ corresponds to a spatial mean. Since the actual local density function is unobserved in its entirety, one estimates $Y$ by taking a sample $s_2$ consisting of $n_2$ points and measuring each of their respective local densities. This sampling procedure is often referred to as \textit{one-phase sampling} (OPS) and $s_2$ is referred to as the terrestrial inventory. In contrast to the one-phase approach, \textit{two-phase} or \textit{double-sampling} procedures use information from two nested samples (phases). Practically speaking, the terrestrial inventory $s_2$ is embedded in a large phase $s_1$ comprising $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest $F$. For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some prediction model. The basic idea of this method is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.

@@

The local density on cluster level $Y_{c}(x)$, which in our case the timber volume per hectare, is then defined as the average of the individual sample plot densities $Y(x_l)$:

@@

\noindent being the empirical model residuals at the cluster level, which by construction of OLS satisfy the important \textit{zero mean residual property}, i.e. $\frac{\sum_{x \in s_{2}} M(x) \hat{R}_{c}(x)}{\sum_{x \in s_{2}} M(x)}=0$.\\

@@

The estimated design-based variance of $ \hat{Y}_{c,G,PSMALL}$ is given by Eq. \ref{eq:psmall_var}. Basically, the first term constitutes the variance introduced by the uncertainty in the regression coefficients, whereas the second term expresses the variance caused by estimating the exact auxiliary mean in $G$ using a non-exhaustive sample $s_{1,G}$. The third term is the variance of the model residuals and thus accounts for the inaccuracies of the model predictions. Note that the first term can also be rewritten using g-weights \cite[pg.14]{mandallaz2016} which ensures some beneficial calibration of the auxiliary variables to the first-phase sample.\\

@@

\begin{equation}\label{eq:error}
error(\hat{Y})_{[\%]}=\frac{\sqrt{\hat{\var}(\hat{Y})}}{\hat{Y}}*100
\end{equation}

We further calculated the 95\% confidence interval for each estimate for visualization purposes. The confidence intervals can also be used heuristically for hypothesis testing to determine whether the point estimates of the three estimators for a given small area are statistically different. The confidence intervals under SRS can be obtained as:

@@

The German federal state Rhineland-Palatinate (\textit{RLP}) is located in the western part of Germany and borders Luxembourg, France and Belgium. With 42.3\% (appr. 8400 km$^2$) of the entire state area (19850 km$^2$) covered by forest, RLP is one of the two states with the highest forest coverage among all federal states of Germany \citep{bwi3}. The forests of RLP are further characterised by a pronounced diversity in bioclimatic growing conditions that have strong influence on the local growth dynamics as well as tree species composition \citep{gauer2005} and are further characterised by large variety of forest structures ranging from characteristic oak coppices (Moselle valley), pure spruce, beech and scots pine forests (i.a. Hunsr{\"u}ck and Palatinate forest) up to mixed forests comprising variable proportions of oak, larch, spruce, Scots pine and beech. Around 82\% of the forest area in RLP are mixed forest stands and 69\% of the forest area exhibit a multi-layered vertical structure. The forest area of RLP are divided into 3 ownership classes, i.e. state forest (27\%), communal forest (46\%) and privately owned forest (27\%). The forest service of RLP has the legal mandate to sustainably manage the state and communal forest area (73\% of the entire forest area), including forest planning, harvesting and the sale of wood \citep{lwaldg_rlp}. For this reason, the entire forest area has been spatially organised in 3 main hierarchical management units (Figure \ref{fig:StudyArea}). On the upper level, RLP has been divided into 45 Forst{\"a}mter (\textit{FA}), which are further divided into a total number of 405 Forstreviere (\textit{FR}). The next level are the forest stands (104'184 in total) for which expert judgements are conducted by SFIs in a 5 to 10 year period in order to set up management strategies for the upcoming 10 years. The FAs and FRs constituted the SA units for which design-based small area estimations of the mean standing timber volume were calculated by incorporating the available terrestrial inventory data of the BWI3 in the estimators described in Section \ref{sec:estimators}. The average area of the SA units were 43'777 ha on the FA-level, and 4624 ha on the FR level.

@@

Rhineland-Palatinate (RLP) is covered by a 2x2 km inventory grid of the German NFI. In the last inventory (BWI3) conducted in the year 2013, timber volume information was derived for 2810 clusters (8092 plots) in the field survey. The local timber volume density on the plot and cluster level for this sample was consequently calculated according to Section \ref{sec:srs_estimator}. In the framework of this survey, the plot center coordinates were re-measured with the differential global satellite navigation system (DGPS) technique. Knowledge about the exact plot positions were considered crucial to provide optimal comparability between the terrestrial observations and the information derived from the auxiliary information. A comparison of the DGPS coordinates with the so-far used target coordinates revealed that 90\% of all horizontal deviations lay in the range of 25 meters. A detailed analysis of horizontal DGPS errors in RLP by \citet{lambrecht2017} indicated that 80\% of the plots should not exceed horizontal DGPS errors of 8 meters. For 162 plots, the DGPS coordinates were replaced by their target coordinates due to missingness or implausible values. The terrestrial sample size $n_{2,G}$ within the FA units was 46 clusters on average and ranged between 11 and 64. Within the FR units, $n_{2,G}$ was considerably smaller with an average of 5 cluster and a range between 0 and 13.

\begin{figure}[H]
	\centering
	\resizebox{0.85\hsize}{!}{\includegraphics*{fig/SAunitsRLP.png}}
	\caption{\textit{Left}: Study Area with delineated FA forest management units. \textit{Right}: Example for each of the three management units (from top to bottom): FA, FR and forest stand unit with overlayed the extended double sampling cluster design. \textit{Green}: Forest stand polygon layer defining the forest area of this study.}
	\label{fig:StudyArea}
\end{figure}

% ----------------------------------------------------------------------- %
\subsection{Extension to double sampling design}
\label{ext_to_2phase}

In order to apply the small area estimators (Section \ref{sec:SAestimators}), the existing NFI design was extended to a double sampling design by densifying the existing systematic 2x2 km grid to a grid size of 500x500 m that constituted the large first phase $s_1$ in accordance to Section \ref{sec:inf_pop} (Figure \ref{fig:StudyArea}, \textit{right}). The existing terrestrial phase $s_2$ was consequently integrated by replacing the target coordinates of the respective $s_1$ clusters by the terrestrially measured DGPS coordinates. For our study, we restricted the \textit{sampling frame} to the communal and state forest. The forest/non-forest decision for each plot was thereby made by a spatial intersection of the plot center coordinates with a polygon layer of the communal and state forest stands provided by the forest service. Using this stand layer provided the advantage to consistently apply the same forest/non-forest definition to the entire sample $s_1$ in order to decide about excluding or including a plot in the sampling frame. The terrestrial sample size $n_2$ was thus reduced to 2055 clusters (5791 plots). Table \ref{tab:fieldata} provides a short descriptive summary about the volume densities and the main attributes of the NFI plots located in the state and communal forest sampling frame. The densification led to an average sample size $n_{1,G}$ of 759  clusters (range: 246 -- 1022) in the FA units, and 88 clusters (range: 1 -- 194) in the FR units.

@@

A prerequisite for the application of the suggested two-phase small area estimators is the identification of suitable auxiliary data available over the entire study area. From 2003 and 2013, the topographic survey institution of RLP conducted an airborne laser scanning acquisition over the entire federal state during leaf-off conditions in order to derive a countrywide digital terrain model (DTM) as well as a digital surface model (DSM). For this study, the recorded LiDAR data was used to create a canopy height model in raster format, providing discrete information about the canopy surface height of the forest area in a spatial resolution of 5 meters (Figure \ref{fig:Auxvars}, \textit{top}). The CHM (Fig. \ref{fig:Auxvars}, \textit{top}) was calculated as the difference between the digital terrain model and the digital surface model that were derived by a Delauney interpolation of the ground and first ALS pulses respectively. A more detailed description of the procedure can be found in \citet{hill2017a}. The CHM provided the most valuable information to be used in the OLS regression model for predicting the timber volume on the plot and cluster level. However, it should be noted that the prolonged acquisition period of the ALS campaign led to the possibility of poor temporal alignment with the BWI3 survey, sometimes up to 10 years. In addition, the quality of the CHM varied substantially as ALS technology evolved over the years.  For example, LiDAR acquisition recorded in 2002 and 2003 exhibited particularly poor quality with about only 0.04 point per m$^2$, whereas more recent datasets contained more than 5 points per m$^2$. Furthermore, CHM information was not available at 16 sample locations due to sensor failures. These plots were deleted from the sampling frame and treated as missing at random. This assumption was considered to be reasonable as the respective sample locations did not exclude specific forest structures.

@@

Additional auxiliary data was derived from a countrywide satellite-based classification map predicting the five main tree species \citep{stoffels2015}, i.e. European beech, Sessile and Pedunculate oak, Norway spruce, Douglas fir and Scots pine (Fig. \ref{fig:Auxvars}, \textit{bottom}). The tree species map has a grid size of 5x5 m and was calculated from 22 bi-temporal satellite images (SPOT5 and RapidEye) using a spatially adaptive classification algorithm \citep{stoffels2012}. As timber volume estimation on the tree level is often based on species-specific biomass and volume equations, the use of tree species information has often been stated as a key factor for improving the precision of timber volume estimates \cite{white2016}. In this respect, incorporating the tree species map was particularly attractive as it predicts five of the seven tree species that are used in the BWI3 taper functions \citep{kublin2013} to calculate the timber volume of a sample tree. However, due to unavailable satellite data, the tree species map excluded one large patch with an area of 415 km$^2$ in the south-west part of RLP covering an entire FA unit consisting of 10 FR units. In 9 additional FR units, the tree species information was also missing for a subset of the sample locations due to two additional patches with areas of 76 km$^2$ and 100 km$^2$ respectively in the northern part of RLP. For these 19 FR units, small area estimation was thus restricted to using only the available CHM information in the regression model. Thus, 411 of 5791 sample locations (approximately 7\%) used to fit the regression model were affected by missing tree species information. A summary of the sample sizes and missing auxiliary data for both the CHM and the tree species map is provided in Table \ref{tab:ssize}.

% neu:
\begin{table}[H]
	\begin{center}
		\caption{Sample size for each phase in entire study area. $n_{\{1,2\},plots}$: number of plots. $n_{\{1,2\}}$: number of clusters. TSPEC: tree species information.}
		\vspace{0.2cm}
		\label{tab:ssize}
		{\small %
			\begin{tabular}{l|r|r|r|r}
				\hline
				\multicolumn{1}{c|}{\textit{Sampling frame}} & \multicolumn{1}{c|}{\textbf{$n_{1,plot}$}}  & \multicolumn{1}{c|}{\textbf{$n_1$}}  & \multicolumn{1}{c|}{\textbf{$n_{2,plot}$}}  & \multicolumn{1}{c}{\textbf{$n_2$}} \\ % <-- added & and content for each column
				\hline \hline
				communal and state forest & 96'854 & 33'365 & 5791 & 2055\\
				\hspace{5mm} \footnotesize missing CHM & \footnotesize 18 & \footnotesize 10 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing TSPEC & \footnotesize 7060  & \footnotesize 3587 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{and} TSPEC & \footnotesize 3 & \footnotesize 2 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{or} TSPEC & \footnotesize 7075 & \footnotesize 3595 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hline \hline
			\end{tabular}
		} %
	\end{center}
\end{table}

@@

The continuous explanatory variables derived from the CHM were the mean canopy height (\textit{meanheight}) and the standard deviation (\textit{stddev}). The quantities were calculated by evaluating the raster values around each sample location within a circle with a predefined radius of 12 meters, i.e. the support. In order to correct for edge effects at the forest border, the intersection of each support area to the state and communal forest area was determined using a polygon mask provided by the state forest service. The percentage of the support within the forest layer was used as the weight $w(x_l)$ introduced in Eq. \ref{eq:Zc(x)} in order to derive the weighted mean of the explanatory variables on the cluster level. Neglecting the calculation of this weight would deteriorate the coherence between explanatory variables computed at the forest boundary and the corresponding local density that already includes a potential boundary adjustment, thus introducing unnecessary noise to the model. The boundary adjustment to the support also makes the sampling frame more consistent for the different data sources (Section \ref{ext_to_2phase}).\par
The ALS acquisition year (ALSyear) was added as a categorical variable in order to account for the time lag with the terrestrial survey as well as to help explain the heterogeneity in the data introduced by the varying ALS quality. In 2008, a sensor error produced particularly poor ALS quality so the year was divided accordingly into two factor levels, denoted \textit{2008\_1} and \textit{2008}. Furthermore, in order to increase the number of observations per factor level the years 2006 and 2007 were pooled together and the same was done for 2012 and 2013.  The result was nine factor levels denoted as \textit{2002}, \textit{2003}, \textit{2007}, \textit{2008\_1}, \textit{2008}, \textit{2009}, \textit{2010}, \textit{2011} and \textit{2012}.

@@

Concerning the existence of outliers or leverage points in the training set for the model, it should be noted that it is more problematic for \psmall{}, \psynth{} and \extpsynth{} to simply remove them as one might be inclined to do in a model-dependent context. Strictly speaking, outlier removal in the design-based context essentially means that those plots, and implicitly any potentially similar plots that were not realized in the selected sample, have been removed from the sampling frame and are no longer considered part of the forest area of interest. While this may be valid for some obvious typos or measurement errors, it is generally not advisable to manipulate the sampling frame after observing data collected from it, especially when the observation in question lies within the small area of interest. However, for sake of completeness, we conducted an analysis of influential observations \citep[pp. 160--167]{fahrmeir2013} on the plot level for the full regression model. We calculated the leverage values and found that 10\% of all observations exceeded a predefined critical threshold, i.e. twice the average of the hat matrix diagonal entries. Further investigation revealed that several leverage points showed unusually large \meanheight{} values compared to their respective timber volume densities. They tended to occur in ALS acquisition years with longer time gaps to the terrestrial survey date and were thus more likely caused by harvesting activities in the sample plot area.  Although these areas likely affected by harvest should clearly not be removed from the sampling frame, it does provided more justification for the inclusion of the \alsyear{} variable to mitigate these effects.

>>>>

The German NFI is a periodic inventory that is carried out every 10 years over the entire forest area of Germany. The most recent inventory (BWI3) was conducted in 2011 and 2012. While information was originally gathered on a systematic 4x4 km grid, some federal states such as Rhineland-Palatinate have switched to a densified 2x2 km grid. The German NFI uses a cluster sampling design, which means that a sample unit consists of at most four sample locations (also referred to as \textit{sample plots}) that are arranged in a square, called \textit{cluster}, with a side length of 150 metres. The number of plots per cluster can vary between 1 and 4 depending on forest/non-forest decisions by the field crews on the individual plot level \citep{bwi3_aufn}. In the field survey of the BWI3, sample trees for timber volume estimation are selected according to the angle count sampling technique \citep{bitterlich1984}, using a basal area factor ($BAF$) of 4 that is respectively adjusted for sample trees at the forest boundary by a geometric intersection of the boundary transect with the individual tree's inclusion circle \citep{bwi3_aufn}. A further inventory threshold for a tree to be recorded is a diameter at breast height ($DBH$) of at least 7 cm. For each sample tree that is selected by this procedure, the DBH, the absolute tree height, the tree diameter at 7 m ($D7$) and the tree species is measured and used to estimate the volume at the tree level. These volume estimates are based on the application of tree species specific taper curves that are adjusted to the set of diameters and corresponding height measurements taken from the respective sample tree \citep{kublin2013}.

@@

The estimators used in this study have been proposed by \citep{mandallaz2013a,mandallaz2013b} and derive their mathematical properties under the so-called infinite population approach. Therefore, we shall first provide a short introduction into this general estimation framework. We start by assuming that the population $P$ of trees $i \in 1,2, ..., N$ within a forest of interest $F$ is exactly defined, and each tree $i$ has a response variable $Y_i$ (e.g. its timber volume) that can be used to define the population mean $Y$ (e.g., the average timber volume per unit area) over $F$. Since a full census of all tree population individuals is almost never feasible, $Y$ has to be estimated based on a sample. In the infinite population approach this sample is a set of points or locations $x$ distributed independently and uniformly over the set of all possible points in $F$. Each point $x$ has an associated local density $Y(x)$ (e.g., the timber volume per unit area) whose spatial distribution is given by a fixed (i.e. non stochastic) piecewise constant function. The population mean $Y$ is mathematically equivalent to the integral of the local density function surface divided by the surface area of $F$, $\lambda(F)$, i.e. $Y=\frac{1}{N} \sum_{i=1}^{N}Y_i=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$, and thus the population mean $Y$ corresponds to a spatial mean. Since the actual local density function is unobserved in its entirety, one estimates $Y$ by taking a sample $s_2$ consisting of $n_2$ points and measuring each of their respective local densities. This sampling procedure is often referred to as \textit{one-phase sampling} (OPS) and $s_2$ is referred to as the terrestrial inventory. In contrast to the one-phase approach, \textit{two-phase} or \textit{double-sampling} procedures use information from two nested samples (phases). Practically speaking, the terrestrial inventory $s_2$ is embedded in a large phase $s_1$ comprising $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest $F$. For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some prediction model. The basic idea of this method is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.

@@

The local density on cluster level $Y_{c}(x)$, which is in our case the timber volume per hectare, is then defined as the average of the individual sample plot densities $Y(x_l)$:

@@

\noindent being the empirical model residuals at the cluster level, which by construction of OLS satisfy the important zero mean residual property, i.e. $\frac{\sum_{x \in s_{2}} M(x) \hat{R}_{c}(x)}{\sum_{x \in s_{2}} M(x)}=0$.\\

@@

The estimated design-based variance of $ \hat{Y}_{c,G,PSMALL}$ is given by Eq. \ref{eq:psmallvar}. Basically, the first term constitutes the variance introduced by the uncertainty in the regression coefficients, whereas the second term expresses the variance caused by estimating the exact auxiliary mean in $G$ using a non-exhaustive sample $s_{1,G}$. The third term is the variance of the model residuals and thus accounts for the inaccuracies of the model predictions. Note that the first term can also be rewritten using g-weights \cite[pg.14]{mandallaz2016} which ensures some beneficial calibration of the auxiliary variables to the first-phase sample.\\

@@

\begin{equation}\label{eq:error}
error_{[\%]}=\frac{\sqrt{\hat{\var}(\hat{Y})}}{\hat{Y}}*100
\end{equation}

We further calculated the 95\% confidence interval for each estimate for visualization purposes. The confidence intervals can also be used heuristically for hypothesis testing to determine whether the point estimates of the three estimators for a given small area are statistically different. The confidence intervals for the SRS estimator can be obtained as:

@@

The German federal state Rhineland-Palatinate (\textit{RLP}) is located in the western part of Germany and borders Luxembourg, France and Belgium. With 42.3\% (appr. 8400 km$^2$) of the entire state area (19850 km$^2$) covered by forest, RLP is one of the two states with the highest forest coverage among all federal states of Germany \citep{bwi3}. The forests of RLP are further characterised by a pronounced diversity in bioclimatic growing conditions that have strong influence on the local growth dynamics as well as tree species composition \citep{gauer2005} and are further characterised by large variety of forest structures ranging from characteristic oak coppices (Moselle valley), pure spruce, beech and scots pine forests (i.a. Hunsr{\"u}ck and Palatinate forest) up to mixed forests comprising variable proportions of oak, larch, spruce, Scots pine and beech. Around 82\% of the forest area in RLP are mixed forest stands and 69\% of the forest area exhibit a multi-layered vertical structure. The forest area of RLP are divided into 3 ownership classes, i.e. state forest (27\%), communal forest (46\%) and privately owned forest (27\%). The forest service of RLP has the legal mandate to sustainably manage the state and communal forest area (73\% of the entire forest area), including forest planning, harvesting and the sale of wood \citep{lwaldg_rlp}. For this reason, the entire forest area has been spatially organised in 3 main hierarchical management units (Figure \ref{fig:StudyArea}). On the upper level, RLP has been divided into 45 Forst{\"a}mter (\textit{FA}), which are further divided into a total number of 405 Forstreviere (\textit{FR}). The next level are the forest stands (104'184 in total) for which expert judgements are conducted by SFIs in a 5 to 10 year period in order to set up management strategies for the upcoming 10 years. The FAs and FRs constituted the SA units for which design-based small area estimations of the mean standing timber volume were calculated by incorporating the available terrestrial inventory data of the BWI3 in the estimators described in Section \ref{sec:estimators}. The average area of the SA units was 43'777 ha on the FA-level, and 4624 ha on the FR level.

@@

Rhineland-Palatinate (RLP) is covered by a 2x2 km inventory grid of the German NFI. In the last inventory (BWI3) conducted in the year 2013, timber volume information was derived for 2810 clusters (8092 plots) in the field survey. The local timber volume density on the plot and cluster level for this sample was consequently calculated according to Section \ref{sec:srs_estimator}. In the framework of this survey, the plot center coordinates were re-measured with the differential global satellite navigation system (DGPS) technique. Knowledge about the exact plot positions were considered crucial to provide optimal comparability between the terrestrial observations and the information derived from the auxiliary information. A comparison of the DGPS coordinates with the so-far used target coordinates revealed that 90\% of all horizontal deviations lay in the range of 25 meters. A detailed analysis of horizontal DGPS errors in RLP by \citet{lambrecht2017} indicated that 80\% of the plots should not exceed horizontal DGPS errors of 8 meters. For 162 plots, the DGPS coordinates were replaced by their target coordinates due to missingness or implausible values. The terrestrial sample size $n_{2,G}$ within the FA units was 46 clusters on average and ranged between 11 and 64. Within the FR units, $n_{2,G}$ was considerably smaller with an average of 5 clusters and a range between 0 and 13.

\begin{figure}[H]
	\centering
	\resizebox{0.85\hsize}{!}{\includegraphics*{fig/SAunitsRLP.png}}
	\caption{\textit{Left}: Study area with delineated FA forest management units. \textit{Right}: Example for each of the three management units (from top to bottom): FA, FR and forest stand unit overlayed with the extended double-sampling cluster design. \textit{Green}: Forest stand polygon layer defining the forest area of this study.}
	\label{fig:StudyArea}
\end{figure}

% ----------------------------------------------------------------------- %
\subsection{Extension to double-sampling design}
\label{ext_to_2phase}

In order to apply the small area estimators (Section \ref{sec:SAestimators}), the existing NFI design was extended to a double-sampling design by densifying the existing systematic 2x2 km grid to a grid size of 500x500 m that constituted the large first phase $s_1$ in accordance to Section \ref{sec:inf_pop} (Figure \ref{fig:StudyArea}, \textit{right}). The existing terrestrial phase $s_2$ was consequently integrated by replacing the target coordinates of the respective $s_1$ clusters by the terrestrially measured DGPS coordinates. For our study, we restricted the sampling frame to the communal and state forest. The forest/non-forest decision for each plot was thereby made by a spatial intersection of the plot center coordinates with a polygon layer of the communal and state forest stands provided by the forest service. Using this stand layer provided the advantage to consistently apply the same forest/non-forest definition to the entire sample $s_1$ in order to decide about excluding or including a plot in the sampling frame. The terrestrial sample size $n_2$ was thus reduced to 2055 clusters (5791 plots). Table \ref{tab:fieldata} provides a short descriptive summary about the volume densities and the main attributes of the NFI plots located in the state and communal forest sampling frame. The densification led to an average sample size $n_{1,G}$ of 759  clusters (range: 246 -- 1022) in the FA units, and 88 clusters (range: 1 -- 194) in the FR units.

@@

A prerequisite for the application of the suggested two-phase small area estimators is the identification of suitable auxiliary data available over the entire study area. From 2003 to 2013, the topographic survey institution of RLP conducted an airborne laserscanning acquisition over the entire federal state during leaf-off conditions in order to derive a countrywide digital terrain model (DTM) as well as a digital surface model (DSM). For this study, the recorded ALS data was used to create a canopy height model (CHM) in raster format, providing discrete information about the canopy surface height of the forest area in a spatial resolution of 5 meters (Fig. \ref{fig:Auxvars}, \textit{top}). The CHM was calculated as the difference between the digital terrain model and the digital surface model that were derived by a Delauney interpolation of the ground and first ALS pulses respectively. A more detailed description of the procedure can be found in \citet{hill2017a}. The CHM provided the most valuable information to be used in the OLS regression model for predicting the timber volume on the plot and cluster level. However, it should be noted that the prolonged acquisition period of the ALS campaign led to the possibility of poor temporal alignment with the BWI3 survey, sometimes up to 10 years. In addition, the quality of the CHM varied substantially as ALS technology evolved over the years.  For example, the ALS acquisitions recorded in 2002 and 2003 exhibited particularly poor quality with about only 0.04 point per m$^2$, whereas more recent datasets contained more than 5 points per m$^2$. Furthermore, CHM information was not available at 16 sample locations due to sensor failures. These plots were deleted from the sampling frame and treated as missing at random. This assumption was considered to be reasonable as the respective sample locations did not exclude specific forest structures.

@@

Additional auxiliary data was derived from a countrywide satellite-based classification map predicting the five main tree species \citep{stoffels2015}, i.e. European beech, Sessile and Pedunculate oak, Norway spruce, Douglas fir and Scots pine (Fig. \ref{fig:Auxvars}, \textit{bottom}). The tree species map has a grid size of 5x5 m and was calculated from 22 bi-temporal satellite images (SPOT5 and RapidEye) using a spatially adaptive classification algorithm \citep{stoffels2012}. As timber volume estimation on the tree level is often based on species-specific biomass and volume equations, the use of tree species information has often been stated as a key factor for improving the precision of timber volume estimates \citep{white2016}. In this respect, incorporating the tree species map was particularly attractive as it predicts five of the seven tree species that are used in the BWI3 taper functions \citep{kublin2013} to calculate the timber volume of a sample tree. However, due to unavailable satellite data, the tree species map excluded one large patch with an area of 415 km$^2$ in the south-west part of RLP covering an entire FA unit consisting of 10 FR units. In 9 additional FR units, the tree species information was also missing for a subset of the sample locations due to two additional patches with areas of 76 km$^2$ and 100 km$^2$ respectively in the northern part of RLP. For these 19 FR units, small area estimation was thus restricted to using only the available CHM information in the regression model. Thus, 411 of 5791 sample locations (approximately 7\%) used to fit the regression model were affected by missing tree species information. A summary of the sample sizes and missing auxiliary data for both the CHM and the tree species map is provided in Table \ref{tab:ssize}.

% neu:
\begin{table}[H]
	\begin{center}
		\caption{Sample size for each phase in entire study area. $n_{\{1,2\},plots}$: number of plots. $n_{\{1,2\}}$: number of clusters. TSPEC: tree species map information.}
		\vspace{0.2cm}
		\label{tab:ssize}
		{\small %
			\begin{tabular}{l|r|r|r|r}
				\hline
				\multicolumn{1}{c|}{\textit{Sampling frame}} & \multicolumn{1}{c|}{\textbf{$n_{1,plot}$}}  & \multicolumn{1}{c|}{\textbf{$n_1$}}  & \multicolumn{1}{c|}{\textbf{$n_{2,plot}$}}  & \multicolumn{1}{c}{\textbf{$n_2$}} \\ % <-- added & and content for each column
				\hline \hline
				communal and state forest & 96'854 & 33'365 & 5791 & 2055\\
				\hspace{5mm} \footnotesize missing CHM & \footnotesize 18 & \footnotesize 10 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing TSPEC & \footnotesize 7060  & \footnotesize 3587 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{and} TSPEC & \footnotesize 3 & \footnotesize 2 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{or} TSPEC & \footnotesize 7075 & \footnotesize 3595 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hline \hline
			\end{tabular}
		} %
	\end{center}
\end{table}

@@

The continuous explanatory variables derived from the CHM were the mean canopy height (\textit{meanheight}) and the standard deviation (\textit{stddev}). The quantities were calculated by evaluating the raster values around each sample location within a circle with a predefined radius of 12 meters, i.e. the support. In order to correct for edge effects at the forest border, the intersection of each support area to the state and communal forest area was determined using a polygon mask provided by the state forest service. The percentage of the support within the forest layer was used as the weight $w(x_l)$ introduced in Eq. \ref{eq:Zc(x)} in order to derive the weighted mean of the explanatory variables on the cluster level. Neglecting the support adjustment would deteriorate the coherence between explanatory variables computed at the forest boundary and the corresponding local density that already includes a potential boundary adjustment, thus introducing unnecessary noise to the model. The boundary adjustment to the support also makes the sampling frame more consistent for the different data sources (Section \ref{ext_to_2phase}).\par
The ALS acquisition year (\textit{ALSyear}) was added as a categorical variable in order to account for the time lag with the terrestrial survey as well as to help explain the heterogeneity in the data introduced by the varying ALS quality. In 2008, a sensor error produced particularly poor ALS quality so the year was divided accordingly into two factor levels, denoted \textit{2008\_1} and \textit{2008}. Furthermore, in order to increase the number of observations per factor level the years 2006 and 2007 were pooled together and the same was done for 2012 and 2013.  The result was nine factor levels denoted as \textit{2002}, \textit{2003}, \textit{2007}, \textit{2008\_1}, \textit{2008}, \textit{2009}, \textit{2010}, \textit{2011} and \textit{2012}.

@@

Concerning the existence of outliers or leverage points in the training set for the model, it should be noted that it is more problematic for \psmall{}, \psynth{} and \extpsynth{} to simply remove them as one might be inclined to do in a model-dependent context. Strictly speaking, outlier removal in the design-based context essentially means that those plots, and implicitly any potentially similar plots that were not realized in the selected sample, have been removed from the sampling frame and are no longer considered part of the forest area of interest. While this may be valid for some obvious typos or measurement errors, it is generally not advisable to manipulate the sampling frame after observing data collected from it, especially when the observation in question lies within the small area of interest. However, for sake of completeness, we conducted an analysis of influential observations \citep[pp. 160--167]{fahrmeir2013} on the plot level for the full regression model. We calculated the leverage values and found that 10\% of all observations exceeding a predefined critical threshold, i.e. twice the average of the hat matrix diagonal entries. Further investigation revealed that several leverage points showed unusually large \meanheight{} values compared to their respective timber volume densities. They tended to occur in ALS acquisition years with longer time gaps to the terrestrial survey date and were thus more likely caused by harvesting activities in the sample plot area.  Although these areas likely affected by harvest should clearly not be removed from the sampling frame, it does provided more justification for the inclusion of the \alsyear{} variable to mitigate these effects.
